Calculating the Solvent Exposed Surface Area of Proteins with Intersecting Spheres

### 2020-06-05 Paper.tex
Removed inconsistent in-text citations.

While atoms are comprised of protons, neutrons, and electrons, electrons are what give them their properties. Electrons make up the volume of an atom, and are more likely to be found closer to the nucleus than away. Because we don’t know for certain where electrons are in an atom, they can be represented by orbitals, or positions that give the probability of electrons being there. Orbitals can take on different shapes, generally resembling some type of dumbbell, but when these orbitals come together, they form a roughly spherical shape. This is why atoms are often modeled as spherical objects. (citation)

@@

In our case, the boundaries of $M$ won't be piecewise geodesic, and thus calculating $\int_{\partial M} k_g ds$ will require knowledge of the circles of intersection between the spheres. With that data calculating $k_g$ simply requires some vector calculus$^1$.
We can calculate the SASA of the entire protein this way, by adding up the contributions of the individual regions $M$ across all atoms.

@@

\begin{comment}
Before launching into how the authors modeled and derived area equations for computational molecular modeling, we should discuss the concept behind $\alpha$-shapes.  $\alpha$-shapes are the foundation of this model, and knowing what they are will help us better understand the derivations presented in the further sections. Hence imagine a set of points in $\R^3$ and we want to describe the 'shape' of these points. One solution is the convex hull of these points, which has been quite well studied. However $\alpha$-shapes are hulls that still include all points, but so so in finer detail than the standard convex hull. Constructing an $\alpha$-shape can be thought of via the following analogy from Fischer$^2$.

@@

\begin{figure}[h!]
\caption{A power diagram comprised of power cells (left) and its Delaunay Triangulation (right).$\lbrack1\rbrack$}
\centerline{\includegraphics[width=3in]{Figure2}}
\end{figure}

@@

\begin{figure}[h!]
\caption{The dual complex (K) of the Delaunay Triangulation only includes those power cells that non-empty intersection.$\lbrack1\rbrack$}
\centerline{\includegraphics[width=3in]{Figure4}}
\end{figure}

@@

\begin{figure}[h!]
\caption{The star of K and link of the star are shown.$\lbrack1\rbrack$}
\centerline{\includegraphics[width=3in]{Figure9}}
\end{figure}

@@

\begin{figure}[h!]
\caption{Two intersecting spheres $S_i$ and $S_j.\lbrack1\rbrack$}
\centerline{\includegraphics[width=3in]{Figure6}}
\end{figure}

@@

\subsection{Distance Preserving Motion}
In distance preserving motion, one sphere rotates about another.
Imagine three spheres interesting at two points. This time, $F$ is going to include all the points contained in all three balls. If the spheres intersect at two points then one sphere contains both of them. Next, consider the caps of the spheres i and k when finding the exposed surface area ($C_{ij}$ and $C_{kj}$). The caps intersect and we can find the parameters for these similar to how we found the formulas for the direction preserving motion. This model says that as you move the two caps apart, keeping their centers in the same position, you will gain an area equal to that of a spherical rectangle.
\begin{figure}[h!]
\caption{Three intersecting spheres at two points: $S_i$, $S_k$ and $S_j.\lbrack1\rbrack$}
\centerline{\includegraphics[width=3in]{Figure7}}
\end{figure}

>>>>

While atoms are comprised of protons, neutrons, and electrons, electrons are what give them their properties. Electrons make up the volume of an atom, and are more likely to be found closer to the nucleus than away. Because we don’t know for certain where electrons are in an atom, they can be represented by orbitals, or positions that give the probability of electrons being there. Orbitals can take on different shapes, generally resembling some type of dumbbell, but when these orbitals come together, they form a roughly spherical shape. This is why atoms are often modeled as spherical objects.

@@

In our case, the boundaries of $M$ won't be piecewise geodesic, and thus calculating $\int_{\partial M} k_g ds$ will require knowledge of the circles of intersection between the spheres. With that data calculating $k_g$ simply requires some vector calculus.
We can calculate the SASA of the entire protein this way, by adding up the contributions of the individual regions $M$ across all atoms.

@@

\begin{comment}
Before launching into how the authors modeled and derived area equations for computational molecular modeling, we should discuss the concept behind $\alpha$-shapes.  $\alpha$-shapes are the foundation of this model, and knowing what they are will help us better understand the derivations presented in the further sections. Hence imagine a set of points in $\R^3$ and we want to describe the 'shape' of these points. One solution is the convex hull of these points, which has been quite well studied. However $\alpha$-shapes are hulls that still include all points, but so so in finer detail than the standard convex hull. Constructing an $\alpha$-shape can be thought of via the following analogy from Fischer.

@@

\begin{figure}[h!]
\caption{A power diagram comprised of power cells (left) and its Delaunay Triangulation (right).}
\centerline{\includegraphics[width=3in]{Figure2}}
\end{figure}

@@

\begin{figure}[h!]
\caption{The dual complex (K) of the Delaunay Triangulation only includes those power cells that non-empty intersection.}
\centerline{\includegraphics[width=3in]{Figure4}}
\end{figure}

@@

\begin{figure}[h!]
\caption{The star of K and link of the star are shown.}
\centerline{\includegraphics[width=3in]{Figure9}}
\end{figure}

@@

\begin{figure}[h!]
\caption{Two intersecting spheres $S_i$ and $S_j$.}
\centerline{\includegraphics[width=3in]{Figure6}}
\end{figure}

@@

\subsection{Distance Preserving Motion}
In distance preserving motion, one sphere rotates about another.
Imagine three spheres interesting at two points. This time, $F$ is going to include all the points contained in all three balls. If the spheres intersect at two points then one sphere contains both of them. Next, consider the caps of the spheres i and k when finding the exposed surface area ($C_{ij}$ and $C_{kj}$). The caps intersect and we can find the parameters for these similar to how we found the formulas for the direction preserving motion. This model says that as you move the two caps apart, keeping their centers in the same position, you will gain an area equal to that of a spherical rectangle.
\begin{figure}[h!]
\caption{Three intersecting spheres at two points: $S_i$, $S_k$ and $S_j$.}
\centerline{\includegraphics[width=3in]{Figure7}}
\end{figure}
